Species persistence under climate and fishing

### 2014-01-29 Writing/MS.tex
more of malin's comments

\documentclass[12pt,english]{article}
\usepackage{geometry}
\geometry{verbose,letterpaper,tmargin=2.54cm,bmargin=2.54cm,lmargin=2.54cm,rmargin=2.54cm}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{natbib}
\bibliographystyle{plainnat}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{brush_style}
\usepackage{setspace}
\usepackage{lineno}
\linenumbers
\title{Species persistence under climate and fishing}
\author{Emma Fuller, Eleanor Brush, Malin Pinsky}
\date{}

@@

\begin{spacing}{1.9}
\begin{flushleft}

@@

When the climate changes, the area in which organisms can survive and reproduce moves through space. 
This change does not occur in isolation but rather appears on a background of other disturbances. We used an 
integrodifference model explicitly accounting for dispersal, reproduction, a shifting environment, and 
harvesting to examine how two disturbances, range shift and harvesting, interact and govern population 
persistence. We found threshold rates of harvesting and of environmental shift that are needed to allow the 
population to persist, studied how these critical parameters depend on the growth rate and dispersal behavior 
of the population, and measured the interactions between the stressors. In particular, we found low but positive 
synergy between the two stressors: harvesting aggravates the population's sensitivity to a shifting range. 
Finally, we introduced two conservation techniques into simulations of the population model -- threshold 
harvest rules and marine protected areas (MPAs) -- and found that 
these approaches could mitigate the negative interaction of the two stressors.  
\hspace{10cm}

\noindent {\bf Keywords:} Climate change, fishing, integrodifference model, synergy, multiple disturbances

@@

There are many stressors that can disturb an ecosystem. Ecologists have quantified the effects of a number of stressors individually \citep{Wilcoveetal1998, Crainetal2008, DarlingCote2008}. However, disturbances rarely occur in isolation and it is therefore important to understand how a system is affected by multiple disturbances together \citep{DoakMorris2010, Fordhametal2013, Foltetal1999}. A perturbation that has little effect when it occurs individually may still amplify the disturbance caused by a coincident perturbation \citep{Crainetal2008, DarlingCote2008}. Synergistic interactions between stressors can be especially problematic if an animal population can persist in the face of a single stressor but the combination of multiple stressors suffices to drive the population extinct (i.e. \citet{Pelletieretal2006}). However it remains difficult to predict when disturbances are likely to interact in ways not anticipated when perturbations are measured individually.

Climate change is a prime example of multiple, possibly interacting, disturbances. Species that are likely to undergo (or already undergoing) shifts in range and phenology are subject to many other disturbances including pollution, habitat fragmentation, over-exploitation, and invasive species \citep{Wilcoveetal1998, Salaetal2000, MEA2005}. A common approach to predicting future population distrubtions under climate change has been to use bioclimatic-envelope models (also known as species distribution models -- SDMs). These statistical models typically correlate presence-absence data with biophysical characteristics such as mean or maximum temperatures, rainfall, or salinity, to explain and predict how species ranges' will differ under climate change \citep{Elithetal2006, GuisanThuiller2005, GuisanZimmerman2000}. Despite these models' widespread adoption, SDMs have frequently been criticized as oversimplified as they lack species interactions, dispersal and reproductive processes \citep{KearneyPorter2009, Zarnetskeetal2012, Robinsonetal2011}. Recent work on range shifts has addressed some of these gaps by explicitly including dispersal and reproduction \citep{Berestyckietal2009, ZhouKot2011}. However these models only address one disturbance: that of climate-driven range shifts. Not only does climate change disturb ecological systems via range shifts (for example), but these range shifts may interact with anthropogenetic stressors already present. Marine fish, for example, are already moving \citep{Perryetal2005, HiddinkHoftstede2008, Rijnsdorpetal2009, Dulvyetal2008, Simpsonetal2011} and projected to move more  \citep{Kelletal2005, Mackenzieetal2007}. These organisms not only are faced with ocean acidification and temperature driven range shifts \citep{Pinskyetal2013, Barryetal1995, Nyeetal2009}, but are often harvested commercially.

Climate change and fishing have been identified as the two largest human impacts on the ocean \citep{Halpernetal2008} so there is a need to understand whether and how these disturbances will interact in the future. Predictions for how climate change will drive shifts in marine systems have focused on how physiological limitations \citep{PortnerKnust2007} and global fish distributions \citep{Cheungetal2008, Cheungetal2009, Hareetal2010, Lenoiretal2011} may change. Work considering the joint impacts of climate and fishing often consider climate fluctuations (large anomalies around the mean) rather than directional changes in climate \citep{WaltersParma1996, KingMcFarlane2006}. Work that does consider the effects of climate-driven range shifts on fishing typically are case-specific and detailed, integrating multiple drivers and disturbances \citep{Cheungetal2010, Lindegrenetal2010, Brownetal2010, Merinoetal2010, Merinoetal2010b, Plaganyietal2011, Ainsworthetal2011, Zhangetal2011, Barangeetal2011, Howardetal2013}. These predicted impacts are important for management and conservation planning \citep{Allisonetal2009}, however these models are so complex that understanding the relative importance of particular drivers, disturbances, and interactions is difficult (but see \citet{Nyeetal2013} for an approach using ecosystem-level models to discern relative importance of disturbances). Here we develop models of intermediate complexity \citep[sensu][]{Gaylordetal2005} that can help clarify the importance of two interacting disturbances: temperature-driven range shifts and harvest. Empirical case studies have identified interactions between overfishing and temperature-driven range shifts that suggest synergy in the magnitude of the disturbance \citep{Lingetal2009} and have demonstrated synergistic effects in micosm experiments between warming temperatures, harvesting and connectivity \citep{Moraetal2007}. This empirical work underscores the importance of understanding how range shifts and harvesting interact and, to our knowledge, there has been limited work examining interactions between harvesting and range shifts.

Here we construct an analytical model of a fish population subject to both harvesting and climate change-induced range-shift. We explicitly include two mechanistic processes central to species' responses to climate and fishing: reproduction and dispersal. Previous work has highlighted the importance of these two processes and their vulnerability to climate change \citep{Fordhametal2013, Hastingsetal2005}. We focus on determining how climate-driven range shifts and fishing interact to affect population persistence. We find that these disturbances interact additively, not multiplicatively which is often the default assumption when considering mortality.

We also examine the effect of marine protected areas (MPAs) on species persistence. Protected areas have been suggested as a key form of climate insurance and stepping stones to help species keep up with a changing environment \citep{Thomasetal2012, Hannahetal2007}. MPAs are frequently recommended for conservation of biodiversity and improved fisheries yield \citep{Gainesetal2010}, and we evaluate whether MPAs established for those purposes could improve species persistence when habitat is shifting rapidly. We find that MPAs can help a species persist with higher harvesting pressure, but does not change the maximum climate velocity with which a species can keep up.

@@

\noindent To make stability easier to analyze, we approximated this kernel, as described in the Appendix.  Analytical results for a separable sinusoidal kernel are also described in the Appendix.  We used 
simulations to analyze a Laplace dispersal kernel that is not amenable to this method, as described below.

For each kernel, the population's ability to persist depends on properties of the population itself: the expected distance a larva disperses ($\langle d \rangle$) and the intrinsic growth rate ($R_0$); properties of the environment: the 
length of the viable patch ($L$) and how quickly the environment is shifting ($c$); and the harvesting rate ($h
$). If the environment shifts more quickly than the critical rate $c^*$ or the population is harvested at more than 
the critical rate $h^*$ then the population will not be able to persist, as described in the Appendix.  For a Gaussian kernel, the critical rates $c^*$ and $h^*$ are those values of $c$ and $h$ such that 
$$R_0(1-h)2\sqrt{2}\exp\left(\frac{-c^2}{8D}\right)\left[\text{erf}\left(\frac{L-c}{2\sqrt{2D}}\right)-\text{erf}\left(\frac{-L-c}{2\sqrt{2D}}\right)\right]=1.$$
A similar expression for a sinusoidal kernel is derived in the appendix.  For both kernels, the critical harvesting proportion can be approximated by a function that looks like 
\begin{equation}
h^*\sim1- \frac{1}{R_0}\cdot C(L)f(\sigma^2,c^2,L^2+3c^2)
\end{equation}
where $C(L,R_0)$ is a decreasing function of the length of the viable patch and the intrinsic growth rate.

\citet{ZhouKot2011} only consider whether a shifting environment will drive a population extinct or not.  To quantify the effects of both a shifting environment and harvesting pressure, we found the total biomass in the equilibrium traveling wave. As in \citet{Latore:1998fk}, for a separable kernel, the equilibrium traveling pulse $n^*(x)$ must satisfy

@@

\noindent Second, we examined harvesting rules more complex than harvesting a constant proportion of the population. Whereas population persistence in the analytical model does not depend on the functional form of recruitment $f$, to perform simulations we must 
specify a recruitment function.  We chose to use a Beverton-Holt function,

@@

\[S = E_\text{hc}-\left(E_\text{h}-E_\text{c}\right).\]

@@

The equilibrium biomass of the population decreases as either the harvesting pressure 
increases or the environmental shifts more quickly (Figure \ref{baseline}). If the equilibrium biomass is $0$, this 
indicates that the harvesting pressure has exceeded the critical harvesting rate $h^*$ and the environmental 
is shifting more quickly than the critical rate of environmental shift $c^*$. As the harvesting rate $h$ increases, 
the critical rate of environmental shift $c^*$ decreases: the environment must move more slowly to 
accommodate the population growing more slowly (Figure \ref{baseline}). Conversely, as the rate of 
environmental shift $c$ increases, the critical harvesting rate $h^*$ decreases (Figure  \ref{baseline}). This 
means that a harvesting rate that is sustainable in the absence of environmental shift may no longer be sustainable if 
the environment starts shifting. The simulations replicate the analytical results with the critical speed $c^*$ 
declining as the critical harvest rate $h^*$ increases and vice versa.

@@

@@

@@

We extended a spatial to study a fish population subject to both the climate-driven range shifts and harvesting. Previous work has highlighted the importance of reproduction and dispersal to the vulnerability of fish populations to climate change \citep[\citet{Fordhametal2013}]{Hastingsetal2005} and our model explicitly includes these processes.  In particular, we derived expressions for critical harvesting rates and critical rates of environmental shift for a separable Gaussian kernel and found these critical rates using numerical simulations of a Laplace kernel.  For each kernel we studied, we found that the higher the growth rate and the better the mean dispersal distance matches the rate of environmental shift, the better a population can adjust to harvest and climate change.  More interestingly, we found a negative relationship between the critical harvesting rate and the rate of environmental shift.  That is, the more quickly the environment shifts the less harvesting it takes to drive the population extinct.  The curved line separating parameters that will allow the population to persist from those that won't is on indication of an interaction between the stressors.

%% discussion of critical rates and curved persistence boundary

@@

The two management strategies we modeled -- harvesting thresholds and MPAs -- both increased the population's biomass at equilibrium and improved its ability to persist.  Protected areas have been advanced as a way to help organisms keep pace with range shifts, as well as to ameliorate anthropogenic disturbances like harvesting and habitat fragmentation\citep{Lawleretal2010, Hannahetal2007,Botsfordetal2001, Gaylordetal2005, HastingsBotsford2003,Thomasetal2012}.  Our results, that spatial management increased the maximum harvesting rate at which the population could survive, corroborate these previous findings.  The management strategies also affect how the two stressors interact with each other.  Specifically, harvesting thresholds can alleviate interactions between the two stressors.   While the management strategies only change harvesting practices and do not directly address affect climate change, understanding how they ameliorate synergistic affects between harvesting and range shifts will help to better implement and place protected areas.

@@

\section{ Acknowledgements}
We'd like to thank our reviewers for providing constructive comments.

\bibliography{fish}

\section{Figures}

\begin{figure}[htbp]
\begin{subfigure}{3in}
\subcaption{\label{biomass}}
\includegraphics[width=\textwidth]{plots/eqbiomass.pdf}
\end{subfigure}
\begin{subfigure}{3in}
\subcaption{\label{rates}}
\includegraphics[width=\textwidth]{plots/critical_rates.pdf}
\end{subfigure}
%\caption{(a) The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis. These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.  (b) The critical harvesting rate on the y-axis as a function of the rate of environmental shift on the x-axis.  Black lines correspond to a growth rate of $R_0=3$, red to $R_0=7$, and blue to $R_0=10$.  Solid lines correspond to an average dispersal distance $\langle d \rangle =0.1$ and dashed lines correspond to an average dispersal distance $\langle d \rangle =0.25$.  These results are from an approximated Gaussian dispersal kernel with $L=1$.}
\label{baseline}
\end{figure}

\pagebreak

Figure \ref{baseline}: (a) The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis. These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.  (b) The critical harvesting rate on the y-axis as a function of the rate of environmental shift on the x-axis.  Black lines correspond to a growth rate of $R_0=3$, red to $R_0=7$, and blue to $R_0=10$.  Solid lines correspond to an average dispersal distance $\langle d \rangle =0.1$ and dashed lines correspond to an average dispersal distance $\langle d \rangle =0.25$.  These results are from an approximated Gaussian dispersal kernel with $L=1$.

\pagebreak

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=3in]{plots/synergy.pdf}
%\caption{Positive synergy between the two stressors.  The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.  These results are from an approximated Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.}
\label{Synergy}
\end{center}
\end{figure}

\pagebreak

Figure \ref{Synergy}: Positive synergy between the two stressors.  The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.  These results are from an approximated Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.

\pagebreak

\begin{subfigure}{.33\textwidth}
\subcaption{}
\includegraphics[width=\textwidth]{plots/eqbiomass_sim.pdf}
\end{subfigure}
\begin{subfigure}{.33\textwidth}
\subcaption{}
\includegraphics[width=\textwidth]{plots/eqbiomass_thresh.pdf}
\end{subfigure}
\begin{subfigure}{.33\textwidth}
\subcaption{}
\includegraphics[width=\textwidth]{plots/eqbiomass_mpa.pdf}
\end{subfigure}
%\caption{The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.  (a) No management.  (b) Threshold harvesting levels.  (c) MPAs.  These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=5$, $K=100$, and $\langle d \rangle =2$. }
\label{management}
\end{figure}

\pagebreak

Figure \ref{management}: The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.  (a) No management.  (b) Threshold harvesting levels.  (c) MPAs.  These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=5$, $K=100$, and $\langle d \rangle =2$.

\section{Appendix}

@@

\end{flushleft}
\end{spacing}

>>>>

\documentclass[]{article}
\usepackage[margin=.75in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{natbib}
\bibliographystyle{plainnat}
\usepackage{todonotes}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{brush_style}
\title{Species persistence under climate and fishing}
\author{Emma Fuller, Eleanor Brush, Malin Pinsky}
\date{}

@@

@@

When the climate changes, the area in which organisms can survive and reproduce moves through space. 
This change does not occur in isolation but rather appears on a background of other disturbances. We used an 
integrodifference model explicitly accounting for dispersal, reproduction, a shifting environment, and 
harvesting to examine how two disturbances, range shift and harvesting, interact and govern population 
persistence. We found threshold rates of harvesting and of environmental shift such that more extreme harvesting or shift will drive the population extinct and studied how these critical parameters depend on the growth rate and dispersal behavior 
of the population.  We then measured the interaction between the stressors and we found low but positive 
synergy between the two stressors: harvesting aggravates the population's sensitivity to a shifting range. 
Finally, we introduced two conservation techniques into simulations of the population model -- threshold 
harvest rules and marine protected areas (MPAs) -- and found that 
these approaches could mitigate the negative interaction of the two stressors.

@@

There are many stressors that can disturb an ecosystem. Ecologists have quantified the effects of a number of stressors individually \citep{Wilcoveetal1998, Crainetal2008, DarlingCote2008}. However, disturbances rarely occur in isolation.  It is therefore important to understand how a system is affected by multiple disturbances together \citep{DoakMorris2010, Fordhametal2013, Foltetal1999}. A perturbation that has little effect when affecting a system individually may amplify the disturbance caused by a coincident perturbation \citep{Crainetal2008, DarlingCote2008}. Synergistic interactions between stressors are especially worrying if an animal population can persist in the face of a single stressor but not in the presence of multiple stressors (e.g.  \citet{Pelletieretal2006}). However, it remains hard to predict when disturbances are likely to interact in ways not predicted by the effects of individual perturbations.

Climate change is a prime example of a stressor occurring in a background of multiple other, possibly interacting, disturbances. These include pollution, habitat fragmentation, invasive species, and over-exploitation \citep{Wilcoveetal1998, Salaetal2000, MEA2005}. It is expected that  species will move their home ranges in response to climate change. Many marine fish are already moving \citep{Perryetal2005, HiddinkHoftstede2008, Rijnsdorpetal2009, Dulvyetal2008, Simpsonetal2011} and are projected to continue to move in the future \citep{Kelletal2005, Mackenzieetal2007}. These species are faced not only with temperature-driven range range shifts but also with anthropogenic stressors including ocean acidification and commercial fishing \citep{Pinskyetal2013, Barryetal1995, Nyeetal2009}. Climate change and fishing have been identified as the two largest human impacts on the ocean \citep{Halpernetal2008} so there is a need to understand whether and how these disturbances will interact in the future.

Many methods have been proposed to help mitigate the effects of harvesting on fish populations. Marine protected areas (MPAs), which provide stepping stones to help species keep up with a changing environment, have been suggested as a key form of climate insurance \citep{Thomasetal2012, Hannahetal2007}. MPAs are frequently recommended for both conservation of biodiversity and improved fisheries yield \citep{Gainesetal2010}. Threshold harvesting levels below which no fish may be harvested have also been proposed as a means of both optimizing yield and conserving resources \citep{Landeetal1995, Landeetal1997}.

A wide range of models have been used to study the effect of climate
change and harvesting on fish populations.  A common approach to predicting future population distributions under climate change has been to use bioclimatic-envelope models (also known as species distribution models -- SDMs). These statistical models have been
used to find correlations between the presence or absence of fish
species and biophysical properties of the environment such as mean or maximum
temperatures, rainfall, or salinity, to explain and predict how species
ranges will differ under climate change \citep{Elithetal2006, GuisanThuiller2005, GuisanZimmerman2000}. Similarly,
projection models of future climate conditions have been used to predict
where conditions will be favorable for fish populations in the future
\citep{Cheungetal2008, Cheungetal2009}.  Despite these models' widespread
adoption, they have  been criticized for being oversimplified
as they lack species interactions, dispersal and reproductive processes
\citep{KearneyPorter2009, Zarnetskeetal2012, Robinsonetal2011}.

Some recent work on range shifts has addressed these gaps by explicitly including dispersal and reproduction \citep{Berestyckietal2009, ZhouKot2011}, which have been highlighted as important factors in fish populations' vulnerability to climate change \citep{Fordhametal2013, Hastingsetal2005}.  These models only address one disturbance, climate-driven range shifts, and do not consider other anthropogenic stressors.  The models of Zhou and Kot
\citep{ZhouKot2011} and Berestycki et al. \citep{Berestyckietal2009} are of
intermediate complexity (sensu
\citep{Gaylordetal2005}).  Work that consider both climate-driven
range shifts and fishing is often case-specific and detailed, including
multiple drivers and disturbances \citep{Cheungetal2010, Lindegrenetal2010, Brownetal2010, Merinoetal2010, Merinoetal2010b, Plaganyietal2011, Ainsworthetal2011, Zhangetal2011,
Barangeetal2011, Howardetal2013}. These cumulative
impacts are important for management and conservation planning
\citep{Allisonetal2009}.  However, these models are
so complex that understanding the relative importance of particular
drivers, disturbances, and interactions is difficult (but see
\citep{Nyeetal2013} for an approach using ecosystem-level models to discern
relative importance of disturbances).

We analyzed an integrodifference model to study the effect of temperature-driven range shifts and
harvesting, two important interacting disturbances, on the persistence of a marine population. In particular, we extended the model of
Zhou and Kot \citep{ZhouKot2011}, which describes a fish population
reproducing and dispersing while their range shifts, by
additionally including harvesting pressure and introducing two different
conservation measures. We found that the disturbances interact
synergistically. We also found that MPAs can help a species persist with
higher harvesting pressure, but they do not change the maximum speed of
habitat shift with which a species can keep up.

@@

\noindent To find analytical results about the Gaussian kernel, we approximated the kernel, as described in the Appendix.  Analytical results for a separable sinusoidal kernel are also described in the Appendix.  We used 
simulations to analyze a Laplace dispersal kernel that is not amenable to this method, as described below.

For each kernel, the population's ability to persist depends on properties of the population itself-- the expected distance a larva disperses ($\langle d \rangle$) and the intrinsic growth rate ($R_0$); properties of the environment-- the 
length of the viable patch ($L$) and how quickly the environment is shifting ($c$); and the harvesting rate ($h
$). If the environment shifts more quickly than the critical rate $c^*$ or the population is harvested at more than 
the critical rate $h^*$ then the population will not be able to persist, as described in the Appendix.  For a Gaussian kernel, the critical rates $c^*$ and $h^*$ are those values of $c$ and $h$ such that 
$$R_0(1-h)2\sqrt{2}\exp\left(\frac{-c^2}{8D}\right)\left[\text{erf}\left(\frac{L-c}{2\sqrt{2D}}\right)-\text{erf}\left(\frac{-L-c}{2\sqrt{2D}}\right)\right]=1.$$
A similar expression for a sinusoidal kernel is derived in the appendix.  For both kernels, the critical harvesting proportion can be approximated by a function that looks like 
\begin{equation}
h^*\sim1- \frac{1}{R_0}\cdot C(L,R_0)f(\sigma^2,c^2,L^2+3c^2)
\end{equation}
where $C(L,R_0)$ is a decreasing function of the length of the viable patch and the intrinsic growth rate.

\citet{ZhouKot2011} only consider whether a shifting environment will drive a population extinct or not.  To quantify the effects of both a shifting environment and harvesting pressure, we additionally found the total biomass in the equilibrium traveling wave.  For a separable kernel, the equilibrium traveling pulse $n^*(x)$ must satisfy \citep{Latore:1998fk}

@@

\noindent Second, we examined harvesting rules more complex than harvesting a constant proportion of the population. Whereas population persistence in the analytical model does not depend on the functional form of recruitment $f$, to perform simulations we must 
specify a recruitment function.  Again, we chose to use a Beverton-Holt function.

@@

\[S = E_\text{hc}-\left(E_\text{h}+E_\text{c}\right).\]

@@

As the harvesting rate $h$ increases, 
the critical rate of environmental shift $c^*$ decreases: the environment must move more slowly to 
accommodate the population growing more slowly (Figure \ref{baseline}). Conversely, as the rate of 
environmental shift $c$ increases, the critical harvesting rate $h^*$ decreases (Figure  \ref{baseline}). This 
means that a harvesting rate that is sustainable in the absence of environmental shift may no longer be sustainable if 
the environment starts shifting.    When the harvesting pressure has exceeded the critical harvesting rate $h^*$ or the environmental 
is shifting more quickly than the critical rate of environmental shift $c^*$, the biomass of the population at equilibrium will be equal to $0$.  Before those thresholds are reached, the equilibrium biomass of the population decreases as either the harvesting pressure 
increases or the environmental shifts more quickly (Figure \ref{baseline}). The simulations replicate the analytical results with the critical speed $c^*$ 
declining as the critical harvest rate $h^*$ increases and vice versa.

\begin{figure}[htbp]
\begin{subfigure}{3in}
\subcaption{\label{rates}}
\includegraphics[width=\textwidth]{plots/critical_rates.pdf}
\end{subfigure}
\begin{subfigure}{3in}
\subcaption{\label{biomass}}
\includegraphics[width=\textwidth]{plots/eqbiomass.pdf}
\end{subfigure}

\caption{(a) The critical harvesting rate on the y-axis as a function of the rate of environmental shift on the x-axis.  Black lines correspond to a growth rate of $R_0=3$, red to $R_0=7$, and blue to $R_0=10$.  Solid lines correspond to an average dispersal distance $\langle d \rangle =0.1$ and dashed lines correspond to an average dispersal distance $\langle d \rangle =0.25$.  (b) The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis. These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.    These results are from an approximated Gaussian dispersal kernel with $L=1$.}
\label{baseline}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=3in]{plots/synergy.pdf}
\caption{Positive synergy between the two stressors.  The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.  These results are from an approximated Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.}
\label{Synergy}
\end{center}
\end{figure}

@@

\begin{subfigure}{.33\textwidth}
\subcaption{}
\includegraphics[width=\textwidth]{plots/eqbiomass_sim.pdf}
\end{subfigure}
\begin{subfigure}{.33\textwidth}
\subcaption{}
\includegraphics[width=\textwidth]{plots/eqbiomass_thresh.pdf}
\end{subfigure}
\begin{subfigure}{.33\textwidth}
\subcaption{}
\includegraphics[width=\textwidth]{plots/eqbiomass_mpa.pdf}
\end{subfigure}
\caption{The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.  (a) No management.  (b) Threshold harvesting levels.  (c) MPAs.  These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=5$, $K=100$, and $\langle d \rangle =2$. }
\label{management}
\end{figure}

@@

We extended a spatial model to study a fish population subject to both the climate-driven range shifts and harvesting. Previous work has highlighted the importance of reproduction and dispersal to the vulnerability of fish populations to climate change \citep[\citet{Fordhametal2013}]{Hastingsetal2005} and our model explicitly includes these processes.  In particular, we derived expressions for critical harvesting rates and critical rates of environmental shift for a separable Gaussian kernel and found these critical rates using numerical simulations of a Laplace kernel.  For each kernel we studied, we found that the higher the growth rate and the better the mean dispersal distance matches the rate of environmental shift, the better a population can adjust to harvest and climate change.  More interestingly, we found a negative relationship between the critical harvesting rate and the rate of environmental shift.  That is, the more quickly the environment shifts the less harvesting it takes to drive the population extinct.  The curved line separating parameters that will allow the population to persist from those that won't is on indication of an interaction between the stressors.

@@

The two management strategies we modeled -- harvesting thresholds and MPAs -- both increased the population's biomass at equilibrium and improved its ability to persist.  Protected areas have been advanced as a way to help organisms keep pace with range shifts, as well as to ameliorate anthropogenic disturbances like harvesting and habitat fragmentation\citep{Lawleretal2010, Hannahetal2007,Botsfordetal2001, Gaylordetal2005, HastingsBotsford2003,Thomasetal2012}.  Our results, that spatial management increased the maximum harvesting rate and maximum climate velocity at which the population could survive, corroborate these previous findings.  The management strategies also affect how the two stressors interact with each other.  Specifically, harvesting thresholds can alleviate interactions between the two stressors.   While the management strategies only change harvesting practices and do not directly address affect climate change, understanding how they ameliorate synergistic affects between harvesting and range shifts will help to better implement and place protected areas.

@@

\bibliography{fish}
\section{Appendix}

@@
